### 2016-05-03 Thesis/DCG_sb-wi-thesis-template/MSpC/MSpC.tex
Initial Commit

\chapter{Morpheus Spectral Counter: A Computational Tool for Label-Free Quantitative Mass Spectrometry using the Morpheus Search Engine}

\section{Summary}
Label-free quantitative MS based on the Normalized Spectral Abundance Factor (NSAF) has emerged as a straightforward and robust method to determine the relative abundance of individual proteins within complex mixtures.  Here, we present Morpheus Spectral Counter (MSpC) as the first computational tool that directly calculates NSAF values from output obtained from Morpheus, a fast, open-source, peptide-MS/MS matching engine compatible with high-resolution accurate-mass instruments.  NSAF has distinct advantages over other MS-based quantification methods, including a higher dynamic range as compared to isobaric tags, no requirement to align and re-extract MS1 peaks, and increased speed.  MSpC features an easy to use graphic user interface that additionally calculates both distributed and unique NSAF values to permit analyses of both protein families and isoforms/proteoforms.  MSpC determinations of protein concentration were linear over several orders of magnitude based on the analysis of several high-mass accuracy datasets either obtained from PRIDE or generated with total cell extracts spiked with purified Arabidopsis 20S proteasomes.  The MSpC software was developed in C\# and is open sourced under a permissive license with the code made available at \url{http://dcgemperline.github.io/Morpheus_SpC/}.

\section{Introduction}

\section{Methods}

\section{Results and Discussion}

\section{Conclusions}

\bibliographystyle{ieeetr}
\bibliography{MSpC}