### 2016-05-03 Thesis/DCG_sb-wi-thesis-template/frontmatter/abstract.tex

\textbf{FIXME:  basically a placeholder; do not believe}

I did some research, read a bunch of papers, published a couple myself, (pick one):
\begin{enumerate}
	\item ran some experiments and made some graphs,
	\item proved some theorems
\end{enumerate}
and now I have a job.  I've assembled this document in the last couple of months so you will let me leave.  Thanks!