### 2016-05-03 Thesis/DCG_sb-wi-thesis-template/frontmatter/acks.tex

\begin{wbepi}{Isaac Asimov (1988)}
Science doesn't purvey absolute truth. Science is a mechanism, a way of trying to improve your knowledge of nature. It's a system for testing your thoughts against the universe and seeing whether they match. This works not just for the ordinary aspects of science, but for all of life.
\end{wbepi}

Acknowledgements go here.