### 2016-05-03 Thesis/DCG_sb-wi-thesis-template/intro/UPS.tex

\chapter{The Ubiquitin 26S Proteasome System}
\section{Ubiquitin Conjugating Machinery}
	\blindtext
	\subsection{E1s}
	\blindtext
	\subsection{E2s}
	\blindtext
	\subsection{E3s}
	\blindtext
	\subsection{DUBS}
\section{The 26S Proteasome}
	\blindtext
	\subsection{The 20S Core Protease}
	\blindtext
	\subsection{The 19S Regulatory Particle}
	\blindtext
		\subsubsection{Regulatory Particle Base}
		\blindtext
		\subsubsection{Regulatory Particle Lid}
		\blindtext
\section{Proteasome Expression}
\blindtext
\section{Proteasome Assembly}
\blindtext
\section{Proteasome Post-Translational Modification}
\blindtext
\section{Proteasome Degredation}
\blindtext
\section{Proteasome Interacting Proteins}
\blindtext