### 2016-07-10 Thesis/DCG_sb-wi-thesis-template/ProteasomePaper/proteasome.tex
additional structuring
structuring

\chapter{Identifying Core Protease and Regulatory Particle Specific Interactors}

\section{Summary}

\section{Introduction}

\section{Experimental Procedures}

\section{Results}

\section{Discussion}

\section{Conclusions}