### 2016-09-19 Thesis/DCG_sb-wi-thesis-template/Proteasome/contam.tex
proteasome additions

\begin{tabular}{@{}lll@{}}
\toprule
\textbf{ATG Identifier} & \textbf{Symbols}   & \textbf{TAIR10 Description}          \\ \midrule
AT1G03880.1             & CRU2, CRB          & cruciferin 2                         \\
AT3G44300.1             & NIT2, AtNIT2       & nitrilase 2                          \\
AT3G44310.3             & NIT1, ATNIT1, NITI & nitrilase 1                          \\
AT3G44310.2             & NIT1, ATNIT1, NITI & nitrilase 1                          \\
AT4G27140.1             & SESA1, AT2S1       & seed storage albumin 1               \\
AT4G27160.1             & AT2S3, SESA3       & seed storage albumin 3               \\
AT4G27170.1             & SESA4, AT2S4       & seed storage albumin 4               \\
AT4G28520.1             & CRU3, CRC          & cruciferin 3                         \\
AT5G44120.3             & CRA1, ATCRA1, CRU1 & RmlC-like cupins superfamily protein \\
AT5G54740.1             & SESA5              & seed storage albumin 5               \\ \bottomrule
\end{tabular}