### 2016-09-19 Thesis/DCG_sb-wi-thesis-template/Proteasome/ipro.tex

\begin{tabular}{@{}lllllll@{}}
\toprule
Protein & AGI Identifier & InterproScan Signature & Annotation                        & E-value  & Start Residue & End Residue \\ \midrule
PBAC1   & AT3G25545      & PF09754                & Proteasome Assembly Chaperone 2*  & 3.4E-06  & 60            & 247         \\
        &                & PTHR37227              & Family Not Named                  & 1.6E-116 & 2             & 267         \\
PBAC2   & AT3G18940      & PF09754                & Proteasome Assembly Chaperone 2   & 1.0E-29  & 18            & 249         \\
        &                & PTHR12970              & Member 5-induced protein 1 clast3 & 4.8E-102 & 3             & 172         \\
PBAC3   & AT5G14710      & PF10178                & Proteasome Assembly Chaperone 3   & 1.6E-26  & 37            & 120         \\
        &                & PTHR31051              & Proteasome Assembly Chaperone 3   & 2.9E-45  & 1             & 124         \\
PBAC4   & AT1G48170      & PF16093                & Proteasome Assembly Chaperone 4   & 1.4E-17  & 47            & 120         \\
        &                & PTHR33559              & Family Not Named                  & 9.5E-56  & 14            & 153         \\
        &                &                        &                                   &          &               &             \\
PAP1    & AT3G07640      & PTHR37221              & Family Not Named                  & 8.1E-109 & 2             & 241         \\ \bottomrule
\end{tabular}