### 2016-09-19 Thesis/DCG_sb-wi-thesis-template/Proteasome/pag1paps.tex

\begin{longtabu} to \textwidth {
|X[2,l]|
X[5,l]|
X[2,l]|
X[2,l]|
X[2,r]|
X[2,r]|
X[2,r]|
|X[2,r]|}
ATG         & TAIR10 Description                                         & Symbol(s) & Proteasome Annotation & Average MS/MS Count* & Average Sequence Coverage (\%) & Fold Change Compared to Control & P-Value  \\
AT3G18940.1 & clast3-related                                             &           & PBAC2                 & 41.5                 & 56.2                           & 2734.5                          & 1.21E-06 \\
AT3G07640.1 & unknown protein                                            &           & PAP1                  & 49.8                 & 59.9                           & 2462.4                          & 6.05E-06 \\
AT3G25545.1 & unknown protein                                            &           & PBAC1                 & 58.2                 & 68.4                           & 2319.3                          & 2.73E-07 \\
AT3G13330.1 & proteasome activating protein 200                          & PA200     & PA200                 & 54.3                 & 30.1                           & 846.7                           & 3.02E-05 \\
AT5G38650.1 & Proteasome maturation factor UMP1                          &           & Ump1b                 & 11.5                 & 67.4                           & 279.2                           & 1.26E-05 \\
AT1G67250.1 & Proteasome maturation factor UMP1                          &           & Ump1a                 & 8.7                  & 56.0                           & 277.0                           & 1.96E-05 \\
AT3G53970.1 & proteasome inhibitor-related                               &           & PTRE1                 & 6.3                  & 29.4                           & 240.6                           & 9.99E-06 \\
AT1G23520.1 & Domain of unknown function (DUF220)                        &           &                       & 8.7                  & 3.0                            & 165.3                           & 1.02E-03 \\
AT1G01320.2 & Tetratricopeptide repeat (TPR)-like superfamily protein    &           &                       & 2.0                  & 1.7                            & 156.0                           & 4.31E-07 \\
AT1G48170.1 & unknown protein                                            &           & PBAC4                 & 19.8                 & 79.5                           & 138.4                           & 5.30E-02 \\
AT5G14710.1 & Contains InterPro:IPR018788                                &           & PBAC3                 & 7.8                  & 39.4                           & 126.3                           & 5.79E-05 \\
AT3G13090.1 & multidrug resistance-associated protein 8                  & ATMRP8    &                       & 0.0                  & 0.5                            & 86.8                            & 1.01E-01 \\
AT2G47360.1 & unknown protein                                            &           &                       & 3.2                  & 2.6                            & 64.5                            & 3.39E-03 \\
AT2G26780.1 & ARM repeat superfamily protein                             &           & ECM29                 & 3.5                  & 9.7                            & 35.6                            & 1.91E-02 \\
AT4G17100.1 & Endoribonuclease XendoU (InterPro:IPR018998)               &           &                       & 0.2                  & 4.4                            & 26.0                            & 6.55E-05 \\
AT5G27960.1 & AGAMOUS-like 90                                            &           &                       & 0.2                  & 6.3                            & 25.1                            & 2.71E-02 \\
AT1G21530.1 & AMP-dependent synthetase and ligase family protein         &           &                       & 0.2                  & 3.1                            & 24.4                            & 1.22E-05 \\
AT1G23780.1 & F-box family protein                                       &           &                       & 2.0                  & 8.6                            & 21.5                            & 1.63E-04 \\
AT3G26020.2 & Protein phosphatase 2A regulatory B subunit family protein &           &                       & 0.8                  & 2.7                            & 14.5                            & 2.53E-04 \\
AT2G17200.1 & ubiquitin family protein                                   & DSK2      &                       & 2.5                  & 9.4                            & 13.7                            & 1.10E-02 \\
ATCG00680.1 & photosystem II reaction center protein B                   & PSBB      &                       & 0.0                  & 2.2                            & 13.4                            & 3.07E-04 \\
AT5G55220.1 & trigger factor type chaperone family protein               &           &                       & 0.2                  & 2.3                            & 13.2                            & 5.56E-03 \\
AT5G13450.1 & delta subunit of Mt ATP synthase                           & ATP5      &                       & 0.0                  & 4.5                            & 8.5                             & 9.55E-04 \\
AT1G79530.1 & glyceraldehyde-3-phosphate dehydrogenase of plastid 1      & GAPCP-1   &                       & 3.3                  & 17.6                           & 7.4                             & 3.56E-04 \\
\end{longtabu}