### 2016-09-20 Thesis/DCG_sb-wi-thesis-template/frontmatter/acks.tex

\begin{wbepi}{Isaac Asimov (1988)}
Science doesn't purvey absolute truth. Science is a mechanism, a way of trying to improve your knowledge of nature. It's a system for testing your thoughts against the universe and seeing whether they match. This works not just for the ordinary aspects of science, but for all of life.
\end{wbepi}
	To my family, especially my wife Erin: I am forever grateful for your love and support, without which I would not be here today. The many adventures that we have already had are just the start of many more yet to come. To my parents, for instilling a curiosity in me that cannot be quenched, and for their unconditional love and support. To my father, Robert: for encouraging me to pursue the computational side of science and to my mother, Terri: for always encouraging me to teach others. To my brother, Aaron, for our many discussions, on science, life, and playing games. To my twin, Stephen, for always reminding me to have fun. To my good friend, Lucas, for all of your guidance and support when I was completely lost in code.

	My scientific colleagues and mentors, you have pushed me to become a better scientist and to always examine the world with a critical eye. Rick Vierstra, thank you for always pushing me, and helping me realize that we can always improve. Patrick Masson, thank you for providing a welcoming home during a time of change.

	I owe much of my mentorship to the Vierstra Lab. Richard Marshall, you are a great friend, and a great mentor. Without your guidance, I would have been lost. To Nick Gladman, for always listening to my crazy ideas, and for being a great friend. To Aaron Lomax, and Robb Stankey for your companionship while drinking coffee.

	I will always look back on my time in Madison fondly. To my friends, Richard, Kurt, Carol, and others: thank you for always being there.